---
title: "Rock data set"
author: 'Benjamin '
date: "2024-02-11"
output:
  html_document: default
  pdf_document: default
---
##collection of the data
```{r }
data<-rock

```
#Exploration of the data
```{r}
str(data)
dim(data)
head(data)
```

##checking for missingvalues
```{r}
table(is.na(data))
##checking for categorical/discrete values
unique(data$perm)
##freq of values
table(data$perm)
##converting the permeability in Milli-Darcies into factors
data$perm<-as.factor(data$perm)
```

#Descriptive Analysis on Numeric variables
```{r}
#measuring on the area of the rock
quantiles<-quantile(data$area,seq(0,1,0.20))
#create breaks
breaks<-quantiles
#histogram plot
hist(data$area,breaks = quantiles, main = "Histogram with Quantiles",xlab = "Area of Rock")

```
```{r}
#measuring on the perimeter fo rock
quantiles_peri<-quantile(data$peri,seq(0,1,0.20))
breaks_peri<-quantiles_peri
hist(data$peri,breaks = quantiles_peri,main = "Histogram with Quantiles", xlab = "Perimeter of Rock")
```


```{r}

#measuring the spread on the shape of the rock
quantiles_shp<-quantile(data$shape,seq(0,1,0.20))
breaks_shp<-quantiles_shp
hist(data$shape, breaks = breaks_shp, main = "Histogram with Quantiles",xlab = "Shape of Rock")


```

##Univariant analysis
```{r}
#area of rock
boxplot(data$area, horizontal = T)
```

#univariant analysis 
```{r}
##perimeter of Rock
boxplot(data$peri, horizontal = T)
```

#univariant analysis on shape 
```{r}
boxplot(data$shape,horizontal = T,main="boxplot on shape")

```

##univariant analysis on the discrete data
```{r}
#install.packages("ggplot2")
library(ggplot2)
ggplot(data,aes(as.factor(perm)))+
  geom_bar()
```
##correlations perm vs area
```{r}
#convert to character
data$perm<-as.character(data$perm)
#convert to numeric
data$perm<-as.numeric(data$perm)
cor_ap<-cor(data$area,data$perm)
plot(area~perm, data = data, main = "Scatter Plot with Trendline", 
     xlab = "Perm", ylab = "Area", pch = 16, col = "blue")
##calculate the slope and intercept
slope_ap<-cor_ap * (sd(data$area)/sd(data$perm))
intercept_ap<-mean(data$area)-slope_ap* mean(data$perm)
abline(a=intercept_ap,b=slope_ap,col="red",lwd=2)

```
##correlations shape vs perm
```{r}
cor_sp<-cor(data$shape,data$perm)
plot(shape~perm,data = data,main="Scatterplot with Trendline", xlab="perm",ylab = "Shape",pch=16,col="blue")
##calculate the slope and intercept
slope_sp<-cor_sp * (sd(data$shape)/sd(data$perm))
intercept_sp<-mean(data$shape)-slope_sp* mean(data$perm)
abline(a=intercept_sp,b=slope_sp,col="red",lwd=2)
```

##correlation peri vs perm
```{r}
cor_pp<-cor(data$peri,data$perm)
plot(peri~perm,data = data,main="Scatterplot with Trendline", xlab="perm",ylab = "Peri",pch=16,col="blue")
##calculate the slope and intercept
slope_pp<-cor_pp * (sd(data$peri)/sd(data$perm))
intercept_pp<-mean(data$peri)-slope_pp* mean(data$perm)
abline(a=intercept_pp,b=slope_pp,col="red",lwd=2)
```

##feature scaling
```{r}
data_sc<-scale(data[,1:3])
summary(data_sc[,3])
View(data_sc)
```

```{r}
perm<-data$perm
data_sc<-cbind(data_sc,data.frame(perm))
```


##linear regression
```{r}
lm_model<-lm(perm~., data = data_sc)
summary(lm_model)

```

```{r}
#install.packages("psych") #assuming you dont have the package
library(psych)
pairs.panels(data_sc[c("area","peri","shape","perm")])
```
# Using ggplot for scatter plot
```{r}
###install.packages("ggplot2")
library(ggplot2)
ggplot(data_sc, aes(x = area, y = peri)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Scatter Plot for Area vs Peri", x = "Area of Rock", y = "Perimeter of Rock")
```


```{r}
library(ggplot2)
ggplot(data_sc, aes(x = area, y = shape)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Scatter Plot for Area vs shape", x = "Area of Rock", y = "shape of Rock")

```


```{r}
library(ggplot2)
ggplot(data_sc, aes(x = shape, y = peri)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Scatter Plot for shape vs Peri", x = "shape of Rock", y = "Perimeter of Rock")
```


```{r}
```


```{r}

```


```{r}
```


```{r}
```


